### 2020-03-13 src/s_lda.Rmd
added start of abstract

---
title: "Fake data"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
```

# AR1 examples
```{r}
f_ar1 <- function(rho, sigma, L){
  x <- vector(length = L)
  x[1] <- rnorm(1, 0, sd=sigma)
  for(i in 2:L)
    x[i] = rho * x[i - 1] + rnorm(1, 0, sd=sigma)
  return(x)
}
plot(f_ar1(1.0, 1, 100), type='l')

f_generate_lower_var_four <- function(var_ratio, rho, sigma, L){
  x <- matrix(nrow = L, ncol = 4)
  for(i in 1:3)
    x[, i] <- f_ar1(rho, sigma, L)
  z <- f_ar1(rho, sigma * sqrt(var_ratio), L)
  x[, 4] <- z
  return(x)
}

f_generate_lower_var_four(1/10, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  ggplot(aes(x=value, fill=as.factor(chain))) +
  geom_histogram(position="identity", alpha=0.3)
```

Test LDA on classifying chains
```{r}
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
require(MASS)
fitter <- lda(formula = chain ~ ., 
         data = training_data, 
         prior = c(1,1,1,1)/4)
plda <- predict(object=fitter, newdata=testing_data)
tibble(predicted=plda$class, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
```

Try Caret
```{r}
library(caret)
library(gbm)
# add in extra column of nonsense values since the 
full_data <- full_data %>% 
  mutate(chain=as.factor(chain)) %>% 
  mutate(ones=rnorm(nrow(full_data)))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid)
# plot(gbmFit1)
plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

# plot predictions as fn of value
testing_data %>% 
  arrange(value) %>% 
  mutate(predictions=predict(object=gbmFit1, .)) %>% 
  ggplot(aes(x=value, y=predictions)) +
  geom_jitter()
```