### 2020-03-14 src/s_lda.Rmd
added 8 schools example

---
title: "Fake data"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
```

# AR1 examples
```{r}
f_ar1 <- function(rho, sigma, L){
  x <- vector(length = L)
  x[1] <- rnorm(1, 0, sd=sigma)
  for(i in 2:L)
    x[i] = rho * x[i - 1] + rnorm(1, 0, sd=sigma)
  return(x)
}
plot(f_ar1(1.0, 1, 100), type='l')

f_generate_lower_var_four <- function(var_ratio, rho, sigma, L){
  x <- matrix(nrow = L, ncol = 4)
  for(i in 1:3)
    x[, i] <- f_ar1(rho, sigma, L)
  z <- f_ar1(rho, sigma * sqrt(var_ratio), L)
  x[, 4] <- z
  return(x)
}

f_generate_lower_var_four(1/10, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  ggplot(aes(x=value, fill=as.factor(chain))) +
  geom_histogram(position="identity", alpha=0.3)
```

Test LDA on classifying chains
```{r}
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
require(MASS)
fitter <- lda(formula = chain ~ ., 
         data = training_data, 
         prior = c(1,1,1,1)/4)
plda <- predict(object=fitter, newdata=testing_data)
tibble(predicted=plda$class, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
```

Try Caret
```{r}
library(caret)
library(gbm)
# add in extra column of nonsense values since the 
full_data <- full_data %>% 
  mutate(chain=as.factor(chain)) %>% 
  mutate(ones=rnorm(nrow(full_data)))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid)
# plot(gbmFit1)
plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

# plot predictions as fn of value
testing_data %>% 
  arrange(value) %>% 
  mutate(predictions=predict(object=gbmFit1, .)) %>% 
  ggplot(aes(x=value, y=predictions)) +
  geom_jitter()
```

Perform 1000 replicates
```{r}
f_replicate <- function(){
  full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)
  rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
  training_data <- full_data[rand_samples, ]
  testing_data <- full_data[-rand_samples, ]

  # add in extra column of nonsense values since the method needs at least 2 vars 
  full_data <- full_data %>% 
    mutate(chain=as.factor(chain)) %>% 
    mutate(ones=rnorm(nrow(full_data)))
  training_data <- full_data[rand_samples, ]
  testing_data <- full_data[-rand_samples, ]

  gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)
  plda <- predict(object=gbmFit1, newdata=testing_data)
  a_accuracy <- 
    tibble(predicted=plda, actual=testing_data$chain) %>%
    mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
    summarise(mean(correct)) %>% 
    pull()
  return(a_accuracy)
}

nreplicates <- 1000
accuracies <- vector(length = nreplicates)
for(i in 1:nreplicates){
  if(i%%100==0)
    print(i)
  accuracies[i] <- f_replicate()
}
```

Plot results
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

g1 <- tibble(accuracy=accuracies/0.25, iteration=seq(1, nreplicates, 1)) %>% 
  ggplot(aes(x=accuracy)) +
  geom_histogram() +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  ggtitle("A.") +
  theme(text = element_text(size=14, colour="black")) +
  xlim(1, NA)

g2 <- testing_data %>% 
  arrange(value) %>% 
  mutate(predictions=predict(object=gbmFit1, .)) %>% 
  ggplot(aes(x=value, y=predictions)) +
  geom_jitter(height = 0.3) +
  xlab("Value") +
  ylab("Chain classification") +
  theme(text = element_text(size=14, colour="black")) +
  coord_flip() +
  ggtitle("B.")
```

Trying probabilistic prediction
```{r}
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
full_data <- full_data %>% 
    mutate(chain=as.factor(chain)) %>% 
    mutate(ones=rnorm(nrow(full_data)))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]

gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

nsim <- 1000
mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/0.25)

f_generate_all_same_var_four <- function(var_ratio, rho, sigma, L){
  x <- matrix(nrow = L, ncol = 4)
  for(i in 1:4)
    x[, i] <- f_ar1(rho, sigma, L)
  return(x)
}

full_data <- f_generate_all_same_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
full_data <- full_data %>% 
    mutate(chain=as.factor(chain)) %>% 
    mutate(ones=rnorm(nrow(full_data)))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]

gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)
plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

mAccuracy1 <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy1[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy1)/0.25)

# saveRDS(tibble(unconverged=colMeans(mAccuracy)/0.25,
#        converged=colMeans(mAccuracy1)/0.25),
#        "../output/ar1_two_replicates.rds")
mAR1 <- readRDS("../output/ar1_two_replicates.rds")
g3 <- 
  mAR1 %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series") +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))

pdf("../output/ar1.pdf", width = 12, height = 6)
multiplot(g1, g2, g3, cols = 3)
dev.off()
```

# Eight schools
```{r}
source("eight_schools.data.R")
eight_schools <- list(J=J, y=y, sigma=sigma)
model_cp <- stan_model("eight_schools_cp.stan")
model_ncp <- stan_model("eight_schools_ncp.stan")
```

Run model
```{r}
fit_cp <- sampling(
  model_cp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)
```

# Break each chain into two then make flattened data
## Centered
```{r}
full_data_all <- rstan::extract(fit_cp, permuted=F)
full_data <- array(dim=c(500, 8, 11))
k <- 1
for(i in 1:4){
  first_half <- full_data_all[1:500, i, ]
  second_half <- full_data_all[501:1000, i, ]
  full_data[, k, ] <- first_half
  k <- k + 1
  full_data[, k, ] <- second_half
  k <- k + 1
}

m_flattened <- matrix(nrow = nrow(full_data) * 8, ncol = 12)
k <- 1
for(i in 1:8){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:11] <- full_data[j, i, ]
    m_flattened[k, 12] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V12) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

nsim <- 1000
mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/0.125)
mAccuracy_centered <- mAccuracy
```

```{r}
full_data_all <- rstan::extract(fit_ncp, permuted=F)
nparams <- 19
full_data <- array(dim=c(500, 8, nparams))
k <- 1
for(i in 1:4){
  first_half <- full_data_all[1:500, i, ]
  second_half <- full_data_all[501:1000, i, ]
  full_data[, k, ] <- first_half
  k <- k + 1
  full_data[, k, ] <- second_half
  k <- k + 1
}

m_flattened <- matrix(nrow = nrow(full_data) * 8, ncol = (nparams+1))
k <- 1
for(i in 1:8){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    m_flattened[k, (nparams+1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V20) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/0.125)
mAccuracy_noncentered <- mAccuracy
```

```{r}
# saveRDS(tibble(centered=colMeans(mAccuracy_centered)/0.125,
#        non_centered=colMeans(mAccuracy_noncentered)/0.125),
#        "../output/eight_schools_replicates.rds")
tibble(centered=colMeans(mAccuracy_centered)/0.125,
       non_centered=colMeans(mAccuracy_noncentered)/0.125) %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))
```

## Trying with original data (i.e. 4 chains)
```{r}
full_data <- rstan::extract(fit_cp, permuted=F)
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = 12)
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:11] <- full_data[j, i, ]
    m_flattened[k, 12] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V12) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

nsim <- 1000
mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
mAccuracy_centered <- mAccuracy

full_data <- rstan::extract(fit_ncp, permuted=F)
k <- 1
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = (nparams+1))
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    m_flattened[k, (nparams+1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V20) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
mAccuracy_noncentered <- mAccuracy
```

```{r}
a_df <- tibble(centered=colMeans(mAccuracy_centered)/(1/n_distinct(training_data$chain)),
       non_centered=colMeans(mAccuracy_noncentered)/(1/n_distinct(training_data$chain)))
saveRDS(a_df, "../output/eight_schools_replicates_4.rds")
a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))
```

Trying original data with index variables
```{r}
full_data <- rstan::extract(fit_cp, permuted=F)
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = 13)
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:11] <- full_data[j, i, ]
    if(j < (nrow(full_data)/2))
      m_flattened[k, 12] <- 0
    else
      m_flattened[k, 12] <- 1
    m_flattened[k, 13] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V13) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

nsim <- 1000
mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
mAccuracy_centered <- mAccuracy

full_data <- rstan::extract(fit_ncp, permuted=F)
k <- 1
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = (nparams+2))
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    if(j < (nrow(full_data)/2))
      m_flattened[k, (nparams+1)] <- 0
    else
      m_flattened[k, (nparams+1)] <- 1
    m_flattened[k, (nparams+2)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V21) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)

plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))
varImp(gbmFit1)

plda <- predict(object=gbmFit1, newdata=testing_data, type = "prob")

mAccuracy <- matrix(nrow = nrow(plda),
                    ncol = nsim)
for(i in 1:nsim){
  for(j in 1:nrow(plda)){
    mAccuracy[j, i] <- if_else(which(rmultinom(1, 1, prob = plda[j, ])==1)==testing_data$chain[j], 1, 0)
  }
}
hist(colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
mAccuracy_noncentered <- mAccuracy
```

```{r}
a_df <- tibble(centered=colMeans(mAccuracy_centered)/(1/n_distinct(training_data$chain)),
       non_centered=colMeans(mAccuracy_noncentered)/(1/n_distinct(training_data$chain)))
saveRDS(a_df, "../output/eight_schools_replicates_4_index.rds")
a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))
```

Three uncertainty distributions
```{r}
a_df <- readRDS("../output/eight_schools_replicates.rds")
g1 <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  xlim(0.78, 1.6) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("A.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = "none")
a_df <- readRDS("../output/eight_schools_replicates_4.rds")
g2 <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlim(0.78, 1.6) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("B.") +
  theme(text = element_text(size=14, colour="black"))
a_df <- readRDS("../output/eight_schools_replicates_4_index.rds")
g3 <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("C.") +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))

pdf("../output/eight_schools.pdf", width = 12, height = 6)
multiplot(g1, g2, cols = 2)
dev.off()
```