### 2020-08-27 src/s_bivariate_normal.Rmd
cleaned up graph and addressed Aki's comments

---
title: "Bivariate normal"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

Generate data
```{r}
rmvrnorm2D <- function(n, mux, muy, sigmax, sigmay, rho){
  return(rmvnorm(n, c(mux, muy),
                 matrix(c(sigmax^2, sigmax * sigmay * rho,
                          sigmax * sigmay * rho, sigmay^2),
                        ncol = 2)))
}
# calling function
n <- 2000
chain1 <- rmvrnorm2D(n, 0, 0, 1, 1, 0.0) %>% 
  as.data.frame() %>% 
  mutate(chain=1,
         iter=seq_along(chain))
chain2 <- rmvrnorm2D(n, 0, 0, 1, 1, 0.0) %>% 
  as.data.frame() %>% 
  mutate(chain=2,
         iter=seq_along(chain))
chain3 <- rmvrnorm2D(n, 0, 0, 1, 1, 0.0) %>% 
  as.data.frame() %>% 
  mutate(chain=3,
         iter=seq_along(chain))
chain4 <- rmvrnorm2D(n, 0, 0, 1, 1, 0.9) %>% 
  as.data.frame() %>% 
  mutate(chain=4,
         iter=seq_along(chain))
```

Determine R* distribution
```{r}
full_samples <- array(dim = c(n, 4, 2))
full_samples[, 1, ] <- as.matrix(chain1[, 1:2])
full_samples[, 2, ] <- as.matrix(chain2[, 1:2])
full_samples[, 3, ] <- as.matrix(chain3[, 1:2])
full_samples[, 4, ] <- as.matrix(chain4[, 1:2])

# Calculate R* dist
r_star_vals <- r_star_dist(full_samples)

g <- 
  qplot(r_star_vals) +
  xlab(TeX("$R*$")) +
  xlim(0.9, NA) +
  geom_vline(xintercept = 1, linetype=2) +
  ylab("Count") +
  theme(text=element_text(size=14, colour="black"),
        axis.text = element_text(colour="black"))
ggsave("../output/bivariate.pdf", g, width = 8, height = 6)
```

Determine Rhat
```{r}
chains_stacked <- rbind.data.frame(chain1, chain2, chain3, chain4) %>% 
  mutate(chain=as.factor(chain))

# Use monitor package functionality -- split chains first
half_iter <- max(chains_stacked$iter) / 2
m_samples <- array(dim = c(half_iter,
                           2*n_distinct(chains_stacked$chain),
                           2))
k <- 1
chains_stacked_first <- chains_stacked %>% 
  filter(iter<=half_iter)
chains_stacked_second <- chains_stacked %>% 
  filter(iter>half_iter)

for(i in 1:half_iter){
  temp <- chains_stacked_first %>% 
    filter(iter==i)
  for(j in 1:4){
    temp1 <- temp %>%
      filter(chain==j)
    m_samples[i, j, ] <- as.numeric(temp1[1, 1:2])
    k <- k + 1
  }
}

for(i in 1:half_iter){
  temp <- chains_stacked_second %>% 
    filter(iter==(i + 1000))
  for(j in 5:8){
    temp1 <- temp %>%
      filter(chain==(j - 4))
    m_samples[i, j, ] <- as.numeric(temp1[1, 1:2])
    k <- k + 1
  }
}

mon <- monitor(m_samples)
```