### 2020-08-27 src/s_trends.Rmd

---
title: "Examining the effect of a trend on R*"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)

f_fit_gbm_predict <- function(r){
  rand_samples <- sample(1:nrow(r), 0.7 * nrow(r))
  training_data <- r[rand_samples, ]
  testing_data <- r[-rand_samples, ]

  gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)
  plda <- predict(object=gbmFit1, newdata=testing_data)
  a_accuracy <- 
    tibble(predicted=plda, actual=testing_data$chain) %>%
    mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
    summarise(mean(correct)) %>% 
    pull()
  return(a_accuracy)
}
```

# Consistent trend across all chains
Repeating analysis from Vehtari et a. appendix B
```{r}
conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  trend = c(0, 0.25, 0.5, 0.75, 1),
  rep = 1:10
)
res <- vector("list", nrow(conds))
chains = 4
for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  r <- array(rnorm(iters * chains), c(iters, chains))
  r <- r + seq(-trend, trend, length.out = iters)
  rs <- as.data.frame(monitor_extra(r))

  # R* calculation
  r1 <- r %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable) %>% 
  mutate(ones=rnorm(length(chain)))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- nrow(r) / 2
  r2 <- cbind(r[1:half_rows, ], r[(half_rows+1):nrow(r), ]) %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable) %>% 
  mutate(ones=rnorm(length(chain)))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, trend, rep, rs, "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res <- bind_rows(res)
saveRDS(res, "../output/experiment_trends_all_dim.rds")
```

```{r}
res <- readRDS("../output/experiment_trends_all_dim.rds")
g <- 
  res %>%
  select(iters, trend, Rstar, Rstar_split) %>% 
  melt(id.vars=c("iters", "trend")) %>% 
  mutate(variable=case_when(variable=="Rstar"~"A. R*",
                            variable=="Rstar_split"~"B. Split-R*")) %>% 
  ggplot(aes(y = value, x = trend)) + 
  geom_point() + 
  geom_jitter() + 
  facet_grid(vars(iters), vars(variable)) + 
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=12, colour="black"),
        strip.text.y = element_text(size=12, colour="black"),
        axis.text.x = element_text(size=10, colour="black"),
        axis.text.y = element_text(size=10, colour="black")) +
  xlab("Trend") +
  ylab("Value")
ggsave("../output/trends_all_dim.pdf", g, width = 8, height = 6)
```

# Trend in a single dimension
```{r}
conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  trend = c(0, 0.25, 0.5, 0.75, 1),
  rep = 1:10,
  dimension = c(1, 5, 10, 20)
)
res <- vector("list", nrow(conds))
chains = 4

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  dimension <- conds[i, "dimension"]
  r <- array(rnorm(iters * chains * dimension), c(iters, chains, dimension))
  r[, , 1] <- r[, , 1] + seq(-trend, trend, length.out = iters)
  rs <- as.data.frame(monitor_extra(r))
  r_2d <- matrix(nrow = iters * chains, ncol=(dimension + 2))
  k <- 1
  for(ii in 1:iters){
    for(j in 1:chains){
      r_2d[k, ] <- c(ii, j, r[ii, j, ])
      k <- k + 1
    }
  }
  r_2d <- r_2d %>%
    as.data.frame()
  colnames(r_2d)[1:2] <- c("iter", "chain")

  # R* calculation
  r1 <- r_2d %>% 
    arrange(chain, iter) %>% 
    select(-iter) %>% 
    mutate(chain=as.factor(chain))
  if(dimension==1)
    r1 <- r1 %>% mutate(ones=rnorm(length(chain)))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- iters / 2
  r_1h <- r_2d %>% 
    filter(iter<=half_rows)
  r_2h <- r_2d %>% 
    filter(iter>half_rows) %>% 
    mutate(chain=chain + chains)
  r2 <- bind_rows(r_1h, r_2h) %>% 
    select(-iter) %>% 
    mutate(chain=as.factor(chain))
  if(dimension==1)
    r2 <- r2 %>% mutate(ones=rnorm(length(chain)))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, trend, rep, dimension, rs, "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res <- bind_rows(res)
saveRDS(res, "../output/experiment_trends_one_dim.rds")
```

```{r}
g <- readRDS("../output/experiment_trends_one_dim.rds") %>% 
  select(trend, iters, rep, dimension, Rstar_split) %>% 
  unique() %>%
  select(-rep) %>% 
  melt(c("trend", "iters", "dimension")) %>% 
  filter(variable=="Rstar_split") %>% 
  ggplot(aes(y = value, x = trend)) + 
  geom_point() + 
  geom_jitter() + 
  facet_grid(vars(iters), vars(dimension)) + 
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=12, colour="black"),
        strip.text.y = element_text(size=12, colour="black"),
        axis.text.x = element_text(size=10, colour="black"),
        axis.text.y = element_text(size=10, colour="black")) +
  xlab("Trend") +
  ylab("Value")
g

ggsave("../output/trends_one_dim.pdf", g, width = 8, height = 6)
```

# Trend in the joint distribution over time
```{r}
library(mvtnorm)
rmvrnorm2D <- function(n, mux, muy, sigmax, sigmay, rho){
  return(rmvnorm(n, c(mux, muy),
                 matrix(c(sigmax^2, sigmax * sigmay * rho,
                          sigmax * sigmay * rho, sigmay^2),
                        ncol = 2)))
}

conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  rep = 1:10,
  rho = c(0, 0.2, 0.4, 0.6, 0.8)
)
res <- vector("list", nrow(conds))
chains = 4

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  rep <- conds[i, "rep"]
  rho <- conds[i, "rho"]
  rho_trend <- seq(-rho, rho, length.out = iters)
  r <- array(dim=c(iters, chains, 2))
  for(j in 1:chains)
    for(jj in 1:iters)
      r[jj, j, ] <- rmvrnorm2D(1, 0, 0, 1, 1, rho_trend[jj])
  # take max as just using the Rhat values
  rs <- as.data.frame(monitor_extra(r)) %>% 
    apply(2, max)

 r_2d <- matrix(nrow = iters * chains, ncol=(2 + 2))
  k <- 1
  for(ii in 1:iters){
    for(j in 1:chains){
      r_2d[k, ] <- c(ii, j, r[ii, j, ])
      k <- k + 1
    }
  }
  r_2d <- r_2d %>%
    as.data.frame()
  colnames(r_2d)[1:2] <- c("iter", "chain")

  # R* calculation
  r1 <- r_2d %>% 
    arrange(chain, iter) %>% 
    select(-iter) %>% 
    mutate(chain=as.factor(chain))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- iters / 2
  r_1h <- r_2d %>% 
    filter(iter<=half_rows)
  r_2h <- r_2d %>% 
    filter(iter>half_rows) %>% 
    mutate(chain=chain + chains)
  r2 <- bind_rows(r_1h, r_2h) %>% 
    select(-iter) %>% 
    mutate(chain=as.factor(chain))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, rep, rho, t(rs), "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res1 <- do.call(rbind, res)
saveRDS(res1, "../output/experiment_trends_joint_distribution.rds")
```

```{r}
g <- readRDS("../output/experiment_trends_joint_distribution.rds") %>%
  as.data.frame() %>% 
  select(iters, rho, rep, zsRhat, Rstar_split) %>% 
  unique() %>%
  select(-rep) %>% 
  melt(c("rho", "iters"))  %>% 
  mutate(variable=case_when(variable=="zsRhat"~"A. Split-Rhat",
                            variable=="Rstar_split"~"B. Split-R*")) %>% 
  ggplot(aes(y = value, x = rho)) + 
  geom_jitter() + 
  facet_grid(vars(iters), vars(variable)) + 
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=12, colour="black"),
        strip.text.y = element_text(size=12, colour="black"),
        axis.text.x = element_text(size=10, colour="black"),
        axis.text.y = element_text(size=10, colour="black")) +
  ylab("Value") +
  xlab(TeX("$\\rho$"))
g

ggsave("../output/trends_joint_distribution.pdf", g, width = 8, height = 6)
```

# Stationary -> non-stationary data
Vary rho of AR1 process from stationary->non-stationary generating process
```{r}
f_ar1 <- function(rho, sigma, L){
  x <- vector(length = L)
  x[1] <- rnorm(1, 0, sd=sigma)
  for(i in 2:L)
    x[i] = rho * x[i - 1] + rnorm(1, 0, sd=sigma)
  return(x)
}

conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  trend = c(0, 0.25, 0.5, 0.75, 1),
  rep = 1:10,
  rho = c(0.8, 0.85, 0.9, 0.95, 0.975, 1)
)
res <- vector("list", nrow(conds))
chains = 4
sigma <- 1

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  rho <- conds[i, "rho"]
  r <- array(rnorm(iters * chains), c(iters, chains))
  for(j in 1:chains)
    r[, j] <- f_ar1(rho, sigma, iters)
  rs <- as.data.frame(monitor_extra(r))

  # R* calculation
  r1 <- r %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(ncol(r1)==2)
    r1 <- r1 %>% mutate(ones=rnorm(length(chain)))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- nrow(r) / 2
  r2 <- cbind(r[1:half_rows, ], r[(half_rows+1):nrow(r), ]) %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(ncol(r2)==2)
    r2 <- r2 %>% mutate(ones=rnorm(length(chain)))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, trend, rep, rho, rs, "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res <- bind_rows(res)
saveRDS(res, "../output/experiment_trends_rho_nearer1.rds")

```

```{r}
res <- readRDS("../output/experiment_trends_rho_nearer1.rds")

ggplot(data = res, aes(x = rho, y = zsRhat)) + 
  geom_jitter() + 
  facet_grid(vars(iters)) + 
  geom_hline(yintercept = 1) + 
  ggtitle('R* without splitting') +
  ylim(0, 6)

g <- ggplot(data = res, aes(x = rho, y = Rstar_split)) + 
  geom_jitter() + 
  facet_grid(vars(iters)) + 
  geom_hline(yintercept = 1) + 
  ylab(TeX("Split-R*")) +
  xlab(TeX("$\\rho$")) +
  ylim(0, 6) +
  theme(strip.text.x = element_text(size=12, colour="black"),
        strip.text.y = element_text(size=12, colour="black"),
        axis.text.x = element_text(size=10, colour="black"),
        axis.text.y = element_text(size=10, colour="black"))
ggsave("../output/trends_ar1.pdf", g, width = 8, height = 6)

b_sum <- res %>%
  group_by(iters, rho) %>% 
  summarise(lower=quantile(Rstar_split, 0.025),
            upper=quantile(Rstar_split, 0.975),
            Rstar_split=median(Rstar_split)) %>% 
  mutate(type="summary")
res1 <- res %>%
  bind_rows(b_sum)

g <- ggplot(data = filter(res1, is.na(type)), aes(x = as.factor(iters), y = Rstar_split, group=rho)) + 
  geom_jitter(width = 0.2, height = 0, colour="grey") + 
  geom_pointrange(data=filter(res1, !is.na(type)), aes(ymin=lower, ymax=upper)) +
  geom_line(data=filter(res1, !is.na(type))) +
  facet_wrap(~rho, scales="free") + 
  geom_hline(yintercept = 1) + 
  ylab(TeX("Split-R*")) +
  xlab("# iterations") +
  theme(strip.text.x = element_text(size=12, colour="black"),
        strip.text.y = element_text(size=12, colour="black"),
        axis.text.x = element_text(size=10, colour="black"),
        axis.text.y = element_text(size=10, colour="black"))

ggsave("../output/trends_ar1_transposed.pdf", g, width = 8, height = 6)
```