### 2020-08-27 src/s_wide_distributions.Rmd

---
title: "MVT Normal: wide"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

## Similar numbers of iterations to target dims
Centered
```{r}
# N <- 250
# A <- readRDS("../output/A_matrix.rds")
# model <- stan_model("mvt_250.stan")
# fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
# saveRDS(fit, "../output/mvt_fit_500.rds")
fit <- readRDS("../output/mvt_fit_500.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star <- r_star_dist(full_data)

a_df <- tibble(samples_cp=r_star)

mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

b_df <- tibble(samples_cp=mon$zfsRhat[1:250])

ha <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk ESS",
                            variable=="tail_ess"~"Tail ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk ESS", "Tail ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ha
aa_df <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat,
       source="Centered")
# ggsave("../output/wide_cp_diagnostics.pdf", ha, width = 12, height = 8)
```

Non-centered
```{r}
# A <- readRDS("../output/A_matrix.rds")
# model <- stan_model("mvt_250_ncp.stan")
# fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
# saveRDS(fit, "../output/mvt_ncp_fit_500.rds")
fit <- readRDS("../output/mvt_ncp_fit_500.rds")
full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star <- r_star_dist(full_data)

a_df <- a_df %>%
  mutate(samples_ncp=r_star)
saveRDS(a_df, "../output/mvt_500_all.rds")
a_df <- readRDS("../output/mvt_500_all.rds")
g <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("Model", labels=c("Centered", "Non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.8, 0.8)) +
  ggtitle("A.")
# ggsave("../output/wide_data.pdf", g, width = 12, height = 8)

mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

h <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk-ESS",
                            variable=="tail_ess"~"Tail-ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk-ESS", "Tail-ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ggsave("../output/wide_ncp_diagnostics.pdf", h, width = 12, height = 8)

aa_df <- aa_df %>%
  bind_rows(tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat,
       source="Non-centered"))

h1 <- aa_df %>% 
  melt(id.vars="source") %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"B. Bulk-ESS",
                            variable=="tail_ess"~"C. Tail-ESS",
                            variable=="rhat"~"A. Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_grid(vars(source), vars(relevel(as.factor(variable), "A. Split-Rhat", "B. Bulk ESS", "C. Tail ESS")), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ggsave("../output/wide_both_diagnostics.pdf", h1, width = 12, height = 8)
```

# Wide dataset - standard normal with 10,000 dimensions
```{r}
# N <- 10000
# A <- diag(nrow=N)
# model <- stan_model("mvt_wide.stan")
# fit <- sampling(model, data=list(N=N), iter=400, chains=4)
# saveRDS(fit, "../output/mvt_wide_400.rds")
fit <- readRDS("../output/mvt_wide_400.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star <- r_star_dist(full_data)
a_df <- tibble(samples_400=r_star)
# saveRDS(a_df, "../output/mvt_wide_400_summary.rds")
mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
```

1,000 samples
```{r}
# N <- 10000
# A <- diag(nrow=N)
# model <- stan_model("mvt_wide.stan")
# fit <- sampling(model, data=list(N=N), iter=1000, chains=4)
# saveRDS(fit, "../output/mvt_wide_1000.rds")

fit <- readRDS("../output/mvt_wide_1000.rds")
full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star1 <- r_star_dist(full_data)

a_df <- a_df %>%
  tibble(samples_1000=r_star1)
saveRDS(a_df, "../output/mvt_wide_summaries.rds")
mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
a_df <- readRDS("../output/mvt_wide_summaries.rds")
a_df <- tibble(samples_400=a_df$.$samples_400, samples_1000=a_df$samples_1000)
g1 <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("# of draws", labels=c("400", "1000")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.8, 0.8)) +
  ggtitle("B.")
ggsave("../output/very_wide_data.pdf", g, width = 12, height = 8)
```

```{r}
pdf("../output/mvt_wide_both.pdf", width = 12, height = 6)
multiplot(g, g1, cols = 2)
dev.off()
```