### 2020-10-09 src/s_ml.Rmd
set up infrastructure for hyperparameter runs and ml comparison

---
title: ML comparison"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

Function to fit any ML model in Caret
```{r}
r_star_ml_variety <- function(x, method_, caretGrid){

  split_chains=T
  training_percent=0.7
  if(split_chains)
    x <- split_data(x)
  nparams <- dim(x)[3]
  nchains <- dim(x)[2]
  niter <- dim(x)[1]
  m_flattened <- matrix(nrow = nrow(x) * nchains,
                        ncol = (nparams + 1))
  k <- 1
  for(i in 1:nchains){
    for(j in 1:nrow(x)){
      m_flattened[k, 1:nparams] <- x[j, i, ]
      m_flattened[k, (nparams + 1)] <- i
      k <- k + 1
    }
  }
  m_flattened <- m_flattened %>% 
    as.data.frame()
  colnames(m_flattened)[nparams + 1] <- "chain"
  r <- m_flattened %>% 
    mutate(chain=as.factor(chain))
  # if only 1 param, add in a column of random noise since gbm requires >1 dims
  if(nparams==1)
    r <- r %>% 
      mutate(V_new=rnorm(nrow(r)))

  rand_samples <- sample(1:nrow(r), training_percent * nrow(r))
  training_data <- r[rand_samples, ]
  testing_data <- r[-rand_samples, ]

  if(method_!="multinom")
      gbmFit <- train(chain ~ ., data = training_data, 
                   trControl = trainControl(method = 'none'),
                   method=method_,
                   tuneGrid = caretGrid)
  else
    gbmFit <- train(chain ~ ., data = training_data, 
                 trControl = trainControl(method = 'none'),
                 method=method_,
                 tuneGrid = caretGrid,
                 MaxNWts=10000
                 )
  plda <- predict(object=gbmFit, newdata=testing_data)
  a_accuracy <- 
    tibble(predicted=plda, actual=testing_data$chain) %>%
    mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
    summarise(mean(correct)) %>% 
    pull()
  return(a_accuracy * n_distinct(training_data$chain))
}

f_replicate_gridded <- function(x, method_, caretGrids_){
  r_stars <- vector(length = nrow(caretGrids_))
  for(i in 1:nrow(caretGrids_))
    r_stars[i] <- r_star_ml_variety(x, method_, caretGrids_[i, ])
  return(r_stars)
}

f_run_all <- function(methods, list_of_caret_grids, f_data_generator){
  x <- f_data_generator()
  r_stars <- vector(length = length(methods))
  for(i in seq_along(r_stars))
    r_stars[i] <- max(f_replicate_gridded(x, methods[i], list_of_caret_grids[[i]]))
  return(tibble(r_star=r_stars, method=methods))
}

f_run_all_replicates <- function(niterations, methods, list_of_caret_grids, f_data_generator){
  for(i in 1:niterations){
    print(i)
    temp_df <- f_run_all(methods, list_of_caret_grids, f_data_generator) %>% mutate(iter=i)
    if(i==1)
      big_df <- temp_df
    else
      big_df <- big_df %>% bind_rows(temp_df)
  }
  return(big_df)
}

tunegrid_gbm <- expand.grid(interaction.depth=3, 
                             n.trees = 50,
                             shrinkage=0.1,
                             n.minobsinnode=10)
tunegrid_rf <- tibble(mtry = 1:2)
tunegrid_knn <- tibble(k = c(5, 10, 15, 20, 40))
tunegrid_svm <- tibble(C = c(0.25, 0.5, 0.75))
tunegrid_multinom <- tibble(decay=c(0.1, 0.2, 0.5, 1))
```

# AR1
```{r}
f_ar1 <- function(rho, sigma, L){
  x <- vector(length = L)
  x[1] <- rnorm(1, 0, sd=sigma)
  for(i in 2:L)
    x[i] = rho * x[i - 1] + rnorm(1, 0, sd=sigma)
  return(x)
}

# Generates three chains with same var; one with a different var
f_generate_lower_var_four <- function(var_ratio, rho, sigma, L){
  x <- matrix(nrow = L, ncol = 4)
  for(i in 1:3)
    x[, i] <- f_ar1(rho, sigma, L)
  z <- f_ar1(rho, sigma * sqrt(var_ratio), L)
  x[, 4] <- z
  return(x)
}

f_data_ar1 <- function(){
  temp <- f_generate_lower_var_four(1/3, 0.3, 1, 1000)
  a_array <- array(dim=c(1000, 4, 1))
  a_array[,,1] <- temp
  return(a_array)
}

temp_df <- f_run_all_replicates(2, c("gbm", "rf", "knn", "svmLinear", "multinom"),
          list(tunegrid_gbm, tunegrid_rf, tunegrid_knn, tunegrid_svm, tunegrid_multinom),
          f_data_ar1)
```

# 8 schools
```{r}
source("eight_schools.data.R")
eight_schools <- list(J=J, y=y, sigma=sigma)
model_cp <- stan_model("eight_schools_cp.stan")
model_ncp <- stan_model("eight_schools_ncp.stan")

f_data_8_schools <- function(){
  fit_ncp <- sampling(
    model_ncp, data = eight_schools,
    iter = 2000, chains = 4, refresh = 0,
    control = list(adapt_delta = 0.95)
    )
  x_ncp <- rstan::extract(fit_ncp, permuted=F)
}

temp_df <- f_run_all_replicates(2, c("gbm", "rf", "knn", "svmLinear", "multinom"),
          list(tunegrid_gbm, tunegrid_rf, tunegrid_knn, tunegrid_svm, tunegrid_multinom),
          f_data_8_schools)
```

# Cauchy
```{r}
f_data_cauchy <- function(){
  fit_nom <- stan(file = 'cauchy_alt_1.stan',
                refresh = 0)
  x <- rstan::extract(fit_nom, permuted=F)
  return(x)
}

temp_df <- f_run_all_replicates(2, c("gbm", "rf", "knn", "svmLinear", "multinom"),
          list(tunegrid_gbm, tunegrid_rf, tunegrid_knn, tunegrid_svm, tunegrid_multinom),
          f_data_cauchy)
```

# Multivariate normal
```{r}

```