### 2020-10-12 src/s_eight_schools.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
source("eight_schools.data.R")
eight_schools <- list(J=J, y=y, sigma=sigma)
model_cp <- stan_model("eight_schools_cp.stan")
model_ncp <- stan_model("eight_schools_ncp.stan")
```

Run models
```{r}
fit_cp <- sampling(
  model_cp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)
```

## Break each chain into two then make flattened data
```{r}
# Centered
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star_dist(x)
mon_cp <- monitor_extra(split_data(x))

# Noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star_dist(x)
mon_ncp <- monitor_extra(split_data(x))

c_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="split")
c1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="split")
```

## Look at GBM fitted to data for variable importance
```{r}
caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
x <- rstan::extract(fit_cp, permuted=F)
x <- split_data(x)
# Reshape data into form for R* estimation
m_flattened <- matrix(nrow = nrow(x) * 8, ncol = 12)
k <- 1
for(i in 1:8){
  for(j in 1:nrow(x)){
    m_flattened[k, 1:11] <- x[j, i, ]
    m_flattened[k, 12] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V12) %>% 
  mutate(chain=as.factor(chain))

full_data <- m_flattened
rand_samples <- sample(1:nrow(full_data), 0.7 * nrow(full_data))
training_data <- full_data[rand_samples, ]
testing_data <- full_data[-rand_samples, ]

# Fit gradient-boosed model
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid, verbose=FALSE)
varImp(gbmFit1)
```

## Trying with original data (i.e. 4 chains)
```{r}
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star_dist(x, split_chains = F)
mon_cp <- monitor_extra(fit_cp)
# noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star_dist(x, split_chains = F)
mon_ncp <- monitor_extra(fit_ncp)

d_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="unsplit")
d1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="unsplit")
```

Plot two R* distributions
```{r}
g1 <- 
  c_df %>% 
  melt(id.vars=c("metric", "simulations")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  xlim(0.78, 2) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("A.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none")
g2 <- 
  d_df %>% 
  melt(id.vars=c("metric", "simulations")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlim(0.78, 2) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("B.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none")
# a_df <- readRDS("../output/eight_schools_replicates_4_index.rds")
```

Look at Rhat distributions
```{r}
e_df <- c1_df %>% 
  mutate(var_num=seq_along(metric)) %>% 
  bind_rows(d1_df %>% mutate(var_num=seq_along(metric)))
g3 <- e_df %>%
  select(-metric) %>% 
  melt(id.vars=c("simulations", "var_num")) %>% 
  pivot_wider(id_cols=c("variable", "var_num"), values_from = "value", names_from = "simulations") %>% 
  ggplot(aes(x=unsplit, y=split, colour=as.factor(variable))) +
  geom_point() +
  scale_x_continuous(limits=c(NA, 1.035)) +
  scale_y_continuous(limits=c(NA, 1.035)) +
  geom_vline(xintercept = 1.01, linetype=2) +
  geom_hline(yintercept = 1.01, linetype=2) +
  geom_abline(linetype=3) +
  scale_color_grey("Series", labels=c("Centered", "Non-centered")) +
  ylab("Split chains") +
  xlab("Original chains") +
  ggtitle("C.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black", size=10),
        legend.position = c(0.72, 0.3))

pdf("../output/fig_10.pdf", width = 12, height = 4)
multiplot(g1, g2, g3, cols = 3)
dev.off()
```

# Plot R* versus quantiles
```{r}
x <- rstan::extract(fit_cp, permuted=F)
g1 <- plot_r_star_quantiles(1, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 2.6)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip()

g2 <- plot_r_star_quantiles(2, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 2.6)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip()

library(ggExtra)
library(gridExtra)
f <- ecdf(as.vector(x[, , 2]))
p <- x[, , 2] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none") +
  xlab("Iteration") +
  ylab(TeX("B. $\\tau$ quantile")) +
  ylim(0, 1)
ga <- ggMarginal(p, type="density", margins="y", groupColour = T,
                 yparams = list(bw=0.05))

f <- ecdf(as.vector(x[, , 1]))
p <- x[, , 1] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Iteration") +
  ylab(TeX("A. $\\mu$ quantile")) +
  ylim(0, 1)
gb <- ggMarginal(p, type="density", margins="y", groupColour = T,
                 yparams = list(bw=0.05))

library(cowplot)

pdf("../output/fig_11.pdf", width = 6, height = 4)
plot_grid(gb, g1, ga, g2, nrow = 2, rel_widths = c(1, 0.4), rel_heights = c(0.9, 1))
dev.off()
```