### 2020-10-22 src/s_cauchy.Rmd
added trends

---
title: "Cauchy example"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(mvtnorm)
library(goftest)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

## Few iterations
Generate split-R* distribution from nominal Cauchy fit
```{r}
fit_nom <- stan(file = 'cauchy_nom.stan', seed = 7878,
                refresh = 0)
x <- rstan::extract(fit_nom, permuted=F)

r_star_nom_gbm <- r_star(x, method="gbm", uncertainty = T)
r_star_nom_rf <- r_star(x, method="rf", uncertainty = T)

mon <- monitor_extra(split_data(x))
a_rhat_df <- tibble(nominal=mon$zfsRhat[1:50])
```

Generate split-R* distribution from alternative Cauchy parameterisation
```{r}
fit_alt1 <- stan(file = 'cauchy_alt_1.stan',
                 seed = 7878, refresh = 0)
full_data <- rstan::extract(fit_alt1, permuted=F)

r_star_alt_gbm <- r_star(x, method="gbm", uncertainty = T)
r_star_alt_rf <- r_star(x, method="rf", uncertainty = T)

mon <- monitor_extra(split_data(full_data))
a_rhat_df <-a_rhat_df %>% 
  mutate(alt=mon$zfsRhat[101:150],
         source="r_hat")
```

Plot split-R* values for each parameterisation
```{r}
a_df <- tibble(nominal_gbm=r_star_nom_gbm,
               nominal_rf=r_star_nom_rf,
               alt_gbm=r_star_alt_gbm,
               alt_rf=r_star_alt_rf,
               source="r_star") %>% 
  bind_rows(a_rhat_df) %>% 
  mutate(source=if_else(source=="r_star", "R*", "split-Rhat"))
# g <-
#   a_df %>% 
#   melt() %>% 
#   ggplot(aes(x=value, fill=as.factor(variable))) +
#   geom_histogram(position="identity", alpha=0.8) +
#   scale_fill_grey("Parameterisation", labels=c("nominal", "alternative")) +
#   xlab("Value") +
#   ylab("Count") +
#   geom_vline(xintercept = 1, linetype=2) +
#   theme(text = element_text(size=14, colour="black"),
#         legend.position = "none") +
#   facet_wrap(~source, scales = "free")
# g
```

## 10,000 iterations
Nominal model with 50000 iterations
```{r}
# fit <- stan(file = 'cauchy_nom.stan', seed = 7878,
#             refresh = 0, iter = 50000)
# saveRDS(fit, "../output/fit_nominal_50000.rds")
fit <- readRDS("../output/fit_nominal_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_nom_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_nom_rf <- r_star(full_data, method="rf", uncertainty = T)
mon_nom <- monitor_extra(split_data(full_data))
```

Alternative model with 50000 iterations
```{r}
# fit <- stan(file = 'cauchy_alt_1.stan', seed = 7878,
#             refresh = 0, iter = 50000)
# saveRDS(fit, "../output/fit_alt_50000.rds")
fit <- readRDS("../output/fit_alt_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

r_star_alt_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_alt_rf <- r_star(full_data, method="rf", uncertainty = T)
mon_alt <- monitor_extra(split_data(full_data))
```

Plot
```{r}
b_df <- tibble(nominal_gbm=r_star_nom_gbm,
               nominal_rf=r_star_nom_rf,
               alt_gbm=r_star_alt_gbm,
               alt_rf=r_star_alt_rf,
               source="r_star") %>% 
  bind_rows(tibble(nominal=mon_nom$zfsRhat[1:50],
                alt=mon_alt$zfsRhat[101:150],
                source="r_hat")) %>% 
  mutate(iter="10,000")

c_df <- a_df %>% mutate(iter="1000") %>% 
  bind_rows(b_df) %>% 
  mutate(source=if_else(source=="r_hat", "split-Rhat", source)) %>% 
  mutate(source=if_else(source=="r_star", "R*", source))
saveRDS(c_df, "cauchy_rhat_rstar.rds")
c_df <- readRDS("cauchy_rhat_rstar.rds")
my_lims <- function(x) { if (max(x) < 1.2) c(0.985, 1.03) else c(0.9, 3)}

g <- c_df %>% 
  melt(id.vars=c("source", "iter")) %>% 
  mutate(source=if_else(source=="split-Rhat", "Split-Rhat", as.character(source))) %>% 
  mutate(iter=fct_relevel(as.factor(iter), "1000", "10,000")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(bins=50, position="identity", alpha=0.8) +
  scale_fill_grey("Parameterisation", labels=c("Nominal", "Alternative")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  facet_wrap(~iter+source, scales="free") +
  scale_x_continuous(limits=my_lims) +
  theme(strip.text = element_text(size=16, colour="black"),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=20, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black"),
        legend.position = "top") +
  xlab("Value")

ggsave("../output/cauchy.pdf", g, width = 12, height = 8)
```

## Look at how various measures indicate convergence with iteration
Compare quantiles
```{r}
fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
r_squared <- function(x, lower=0.001){
  qs <- seq(lower, 1-lower, lower)
  q_actual <- map_dbl(qs, ~qcauchy(.))
  r_vals <- vector(length = 50)
  for(i in seq_along(r_vals)){
    q_approx <- quantile(x[, , i], qs)
    aLM <- lm(data=tibble(q_approx, q_actual), formula = q_actual~q_approx)
    a_sum <- summary(aLM)
    r_vals[i] <- a_sum$r.squared
  }
  return(r_vals)
}

log_last <- log10(25000)
log_first <- log10(100)
iter <- round(10^(seq(log_first, log_last, length.out = 20)))
r_vals1 <- map_dbl(iter, ~median(r_squared(x[1:., , ])))
r_vals1_lower <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.025))
r_vals1_upper <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.975))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
r_vals <- map_dbl(iter, ~median(r_squared(x[1:., , ])))
r_vals_lower <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.025))
r_vals_upper <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.975))
```

Compare nominal and alt parameterisations
```{r}
b_df <- 
  tibble(iteration=iter,
         middle=r_vals,
         lower=r_vals_lower,
         upper=r_vals_upper,
         variable="alt") %>% 
  bind_rows(tibble(iteration=iter,
         middle=r_vals1,
         lower=r_vals1_lower,
         upper=r_vals1_upper,
         variable="nom"))
# saveRDS(b_df, "../output/cauchy_r_squared.rds")
b_df <- readRDS("../output/cauchy_r_squared.rds")
g1 <- b_df %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=middle,colour=as.factor(variable))) +
  geom_line(aes(y=middle,colour=as.factor(variable))) +
  ylim(NA, 1) +
  ylab(TeX("$R^2$")) +
  scale_color_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  scale_x_log10() +
  ggtitle(TeX("A. Quantile-$R^2$"))
```

Cramer Von Mises -- seems very noisy
```{r}
library(goftest)
f_cvm <- function(x){
  temp <- cvm.test(x, null = "pcauchy")
  return(temp$p.value)
}

f_cvm_all_dims <- function(x, max_iter){
  vals <- map_dbl(seq(1, 50, 1), ~f_cvm(x[1:max_iter, , .]))
  return(vals)
}

fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
cvm_min <- map_dbl(iter, ~mean(f_cvm_all_dims(x, .)))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
cvm_min <- map_dbl(iter, ~mean(f_cvm_all_dims(x, .)))
```

Anderson-Darling test -- again, noisy
```{r}
library(goftest)
f_ad <- function(x){
  temp <- ad.test(x, null = "pcauchy")
  return(temp$p.value)
}

f_ad_all_dims <- function(x, max_iter){
  vals <- map_dbl(seq(1, 50, 1), ~f_ad(x[1:max_iter, , .]))
  return(vals)
}

fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
cvm_min <- map_dbl(iter, ~mean(f_ad_all_dims(x, .)<0.05))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x", inc_warmup=T)
cvm_min <- map_dbl(iter, ~mean(f_ad_all_dims(x, .)<0.05))
```

Progression of R*
```{r}
fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

library(furrr)
future::plan(multiprocess)
laccuracy <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

laccuracy1 <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

a_df <- 
  tibble(iteration=iter,
         alt=laccuracy1,
       nom=laccuracy)
saveRDS(a_df, "../output/cauchy_r_star_rf.rds")

fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

laccuracy <- future_map_dbl(iter, function(i) r_star(x[1:i, , ], method="gbm"))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

laccuracy1 <- future_map_dbl(iter, function(i) r_star(x[1:i, , ], method="gbm"))

a_df <- 
  tibble(iteration=iter,
         alt=laccuracy1,
       nom=laccuracy)
saveRDS(a_df, "../output/cauchy_r_star_gbm.rds")
```

Add in GBM R*
```{r}
a_df <- readRDS("../output/cauchy_r_star_gbm.rds") %>% 
  mutate(model="GBM")
a_df1 <- readRDS("../output/cauchy_r_star_rf.rds") %>% 
  mutate(model="RF")
g2 <- a_df %>%
  bind_rows(a_df1) %>% 
  melt(id.vars=c("iteration", "model")) %>% 
  ggplot(aes(x=iteration, y=value, colour=as.factor(variable), linetype=as.factor(model))) +
  geom_point() +
  geom_line() +
  ylab(TeX("$R^*$")) +
  scale_color_grey(guide="none") +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.83, 0.77),
        legend.text = element_text(size=12, colour="black"),
        legend.title = element_text(size=14, colour="black"),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.8, NA), breaks = seq(1, 7, 1)) +
  ggtitle(TeX("D. Split-$R^*$"))+
  scale_linetype_discrete("Classifier") +
  guides(
         linetype=guide_legend(nrow = 1, byrow = T))
```

Progression of R* with ranked samples instead
```{r}
fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
future::plan(multiprocess)
unranked <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

x <- z_scale(x)
ranked <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))
cc_df <- tibble(ranked, unranked, iter, source="nominal")

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
future::plan(multiprocess)
unranked <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

x <- z_scale(x)
ranked <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

cc_df <- cc_df %>% 
  bind_rows(tibble(ranked, unranked, iter, source="alt"))
saveRDS(cc_df, "../output/cauchy_ranked_unranked.rds")
cc_df <- readRDS("../output/cauchy_ranked_unranked.rds")

g <- cc_df %>% 
  mutate(source=if_else(source=="alt", "alternative", "nominal")) %>% 
  melt(id.vars=c("iter", "source")) %>% 
  mutate(variable=as.factor(variable)) %>% 
  ggplot(aes(x=iter, y=value, colour=variable)) +
  geom_line() +
  facet_wrap(~source) +
  scale_x_log10() +
  scale_color_grey("Method") +
  xlab("Draws") +
  ylab("R*")
ggsave("../output/cauchy_ranked_unranked.pdf", g)
```

Rhat
```{r}
r_hat_all_params <- function(x){
 nparams <- dim(x)[3] 
 rhat_vals <- map_dbl(seq(1, nparams, 1), ~rhat(split_chains(z_scale(x[ , , .]))))
 return(rhat_vals)
}

fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
future::plan(multiprocess)
iter <- a_df$iteration %>% unique()
r_hat <- future_map_dbl(iter, ~max(r_hat_all_params(x[1:., , ])))
r_hat_lower <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.025))
r_hat_upper <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.975))

fit1 <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
r_hat1 <- future_map_dbl(iter, ~max(r_hat_all_params(x[1:., , ])))
r_hat1_lower <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.025))
r_hat1_upper <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.975))

c_df <- 
  tibble(iteration=iter,
         middle=r_hat1,
         lower=r_hat1_lower,
         upper=r_hat1_upper,
         variable="alt") %>% 
  bind_rows(tibble(iteration=iter,
         middle=r_hat,
         lower=r_hat_lower,
         upper=r_hat_upper,
         variable="nom"))
saveRDS(c_df, "../output/cauchy_r_hat.rds")
c_df <- readRDS("../output/cauchy_r_hat.rds")
g3 <- 
  c_df %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=middle, colour=as.factor(variable))) +
  geom_line(aes(y=middle, colour=as.factor(variable))) +
  ylab(TeX("$\\hat{R}$")) +
  scale_color_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.8, 0.6),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.99, NA), breaks = seq(1, 6, 0.2)) +
  ggtitle(TeX("B. Max split-$\\hat{R}$"))
g3
```

Multivariate Rhat
```{r}
fit1 <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
Rhat_alt <- map_dbl(iter, ~r_hat_multivariate(x[1:., , ]))

fit1 <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
Rhat_nom <- map_dbl(iter, ~r_hat_multivariate(x[1:., , ]))
g_df <- tibble(iteration=iter,
       alt=Rhat_alt,
       nom=Rhat_nom) 
g4 <- 
  tibble(iteration=iter,
       alt=Rhat_alt,
       nom=Rhat_nom) %>% 
  melt(id.vars="iteration") %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=value, colour=as.factor(variable))) +
  geom_line(aes(y=value, colour=as.factor(variable))) +
  ylab(TeX("$\\hat{R}$")) +
  scale_color_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  scale_fill_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.99, NA), breaks = seq(1, 6, 0.2)) +
  ggtitle(TeX("C. Multivariate split-$\\hat{R}$"))
g4
```

Tail ESS
```{r}
fit1 <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
tail_ess_alt <- map_dbl(iter, function(i) min(map_dbl(seq(1, 50, 1), ~ess_tail(x[1:i, , .]))))
bulk_ess_alt <- map_dbl(iter, function(i) min(map_dbl(seq(1, 50, 1), ~ess_bulk(x[1:i, , .]))))

fit1 <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
tail_ess_nom <- map_dbl(iter, function(i) min(map_dbl(seq(1, 50, 1), ~ess_tail(x[1:i, , .]))))
bulk_ess_nom <- map_dbl(iter, function(i) min(map_dbl(seq(1, 50, 1), ~ess_bulk(x[1:i, , .]))))

e_df <- tibble(iteration=iter,
               alt=tail_ess_alt,
               nom=tail_ess_nom)
saveRDS(e_df, "../output/cauchy_tail_ess.rds")
f_df <- tibble(iteration=iter,
               alt=bulk_ess_alt,
               nom=bulk_ess_nom)
saveRDS(f_df, "../output/cauchy_bulk_ess.rds")
e_df <- readRDS("../output/cauchy_tail_ess.rds")
f_df <- readRDS("../output/cauchy_bulk_ess.rds")
g5 <- 
  e_df %>% 
  melt(id.vars="iteration") %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=value, colour=as.factor(variable))) +
  geom_line(aes(y=value, colour=as.factor(variable))) +
  scale_color_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  scale_fill_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 400, linetype=2) +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle(TeX("E. Tail-ESS"))

g6 <- 
  f_df %>% 
  melt(id.vars="iteration") %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=value, colour=as.factor(variable))) +
  geom_line(aes(y=value, colour=as.factor(variable))) +
  scale_color_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  scale_fill_grey("Parameterisation", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 400, linetype=2) +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle(TeX("F. Bulk-ESS"))
```

Combine
```{r}
pdf("../output/cauchy_convergence.pdf", width = 12, height = 8)
multiplot(g1, g4, g5, g3, g2, g6, cols = 2)
dev.off()
```