### 2020-11-09 src/s_ml.Rmd

---
title: ML comparison"
output: html_notebook
---

Note that, to generate the results in this section, we used the shell file 
s_jobScheduler_ml_hypers_set.sh

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

# AR1
```{r}
orders <- c("rf", "xgbTree", "gbm", "knn", "multinom", "svmLinear")
lookup <- tibble(method=orders,
                 new=c("RF", "xgbTree", "GBM", "KNN", "GLM", "SVM"))

temp_df <- readRDS("../data/ml_comp_hypers_set_ar1.rds")  %>% 
  filter(method!="xgbTree") %>%
  select(method, iter, r_star, time) %>% 
  left_join(lookup, by="method") %>% 
  select(-method) %>% 
  rename(method=new) %>% 
  mutate(method=as.factor(method)) %>% 
  mutate(method=fct_relevel(method, lookup$new)) %>% 
  rename(`R*`=r_star, `Time`=time) %>% 
  melt(id.vars=c("method", "iter")) %>% 
  group_by(method, variable) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(model="A. AR(1)")
```

# 8 schools
```{r}
temp_df1 <- readRDS("../data/ml_comp_hypers_set_8_schools.rds")  %>% 
  filter(method!="xgbTree") %>%
  select(method, iter, r_star, time) %>% 
  left_join(lookup, by="method") %>% 
  select(-method) %>% 
  rename(method=new) %>% 
  mutate(method=as.factor(method)) %>% 
  mutate(method=fct_relevel(method, lookup$new)) %>% 
  rename(`R*`=r_star, `Time`=time) %>% 
  melt(id.vars=c("method", "iter")) %>% 
  group_by(method, variable) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(model="D. 8 schools")
```

# Cauchy
```{r}
temp_df2 <- readRDS("../data/ml_comp_hypers_set_cauchy.rds") %>% 
  filter(method!="xgbTree") %>%
  select(method, iter, r_star, time) %>% 
  left_join(lookup, by="method") %>% 
  select(-method) %>% 
  rename(method=new) %>% 
  mutate(method=as.factor(method)) %>% 
  mutate(method=fct_relevel(method, lookup$new)) %>% 
  rename(`R*`=r_star, `Time`=time) %>% 
  melt(id.vars=c("method", "iter")) %>% 
  group_by(method, variable) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(model="C. Cauchy")
```

# Multivariate normal
```{r}
temp_df3 <- readRDS("../data/ml_comp_hypers_set_normal_250.rds") %>% 
  filter(method!="xgbTree") %>%
  select(method, iter, r_star, time) %>% 
  left_join(lookup, by="method") %>% 
  select(-method) %>% 
  rename(method=new) %>% 
  mutate(method=as.factor(method)) %>% 
  mutate(method=fct_relevel(method, lookup$new)) %>% 
  rename(`R*`=r_star, `Time`=time) %>% 
  melt(id.vars=c("method", "iter")) %>% 
  group_by(method, variable) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(model="B. 250D normal")

g1 <- temp_df %>%
  bind_rows(temp_df1) %>% 
  bind_rows(temp_df2) %>% 
  bind_rows(temp_df3) %>% 
  mutate(model=fct_relevel(as.factor(model), "A. AR(1)", "B. 250D normal", "C. Cauchy", "D. 8 schools")) %>%
  ggplot(aes(x=as.factor(method), y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper)) +
  xlab("Method") +
  ylab("Value") +
  theme(text=element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.83, 0.79),
        legend.text = element_text(size=12, colour="black"),
        legend.title = element_text(size=14, colour="black")) +
  facet_grid(vars(variable), vars(model), scales="free")
ggsave("../output/ml_comparison_all.pdf", g1, width = 12, height = 6)
```