### 2020-11-09 src/s_wide_distributions.Rmd

---
title: "MVT Normal: wide"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

## Similar numbers of iterations to target dims
Centered
```{r}
# N <- 250
# A <- readRDS("../output/A_matrix.rds")
# model <- stan_model("mvt_250.stan")
# fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
# saveRDS(fit, "../output/mvt_fit_500.rds")
fit <- readRDS("../output/mvt_fit_500.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)

## R* calcs
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")

a_df <- tibble(samples_cp_gbm=r_star_gbm,
               samples_cp_rf=r_star_rf)

## Rhat cals and ESSs
mon <- monitor_extra(full_data)

f_rhat_max <- function(mon) {
  rhat <- mon$zfsRhat
  rhat1 <- mon$zsRhat
  rhat <- tibble(rhat, rhat1) %>% 
    apply(1, max)
  return(rhat)
}

rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

## plot traditional
ha <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk ESS",
                            variable=="tail_ess"~"Tail ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk ESS", "Tail ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ha
aa_df <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats,
       source="Centered")
# ggsave("../output/wide_cp_diagnostics.pdf", ha, width = 12, height = 8)
```

Non-centered
```{r}
# A <- readRDS("../output/A_matrix.rds")
# model <- stan_model("mvt_250_ncp.stan")
# fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
# saveRDS(fit, "../output/mvt_ncp_fit_500.rds")
fit <- readRDS("../output/mvt_ncp_fit_500.rds")
full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")

a_df <- a_df %>%
  mutate(samples_ncp_gbm=r_star_gbm,
         samples_ncp_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_500_all.rds")
a_df <- readRDS("../output/mvt_500_all.rds")
a_df_gbm <- a_df %>%
  select(contains("gbm")) %>% 
  rename(samples_cp=samples_cp_gbm,
         samples_ncp=samples_ncp_gbm) %>% 
  mutate(Classifier="A. GBM")
a_df_rf <- a_df %>%
  select(contains("rf")) %>% 
  rename(samples_cp=samples_cp_rf,
         samples_ncp=samples_ncp_rf) %>% 
  mutate(Classifier="B. RF")

g <- 
  a_df_gbm %>%
  bind_rows(a_df_rf) %>% 
  melt(id.vars="Classifier") %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("Model", labels=c("Centered", "Non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=20, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.45, 0.5)) +
  facet_wrap(~Classifier)
# ggsave("../output/wide_data.pdf", g, width = 12, height = 8)

mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

h <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk-ESS",
                            variable=="tail_ess"~"Tail-ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk-ESS", "Tail-ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
# ggsave("../output/wide_ncp_diagnostics.pdf", h, width = 12, height = 8)

## plot all traditional measures
aa_df <- aa_df %>%
  bind_rows(tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats,
       source="Non-centered"))

h1 <- aa_df %>% 
  melt(id.vars="source") %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"B. Bulk-ESS",
                            variable=="tail_ess"~"C. Tail-ESS",
                            variable=="rhat"~"A. Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_grid(vars(source), vars(relevel(as.factor(variable), "A. Split-Rhat", "B. Bulk ESS", "C. Tail ESS")), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=18, colour="black"),
        axis.text=element_text(size=18, colour="black"),
        strip.text = element_text(size=18, colour="black"),
        axis.title.x = element_blank())
ggsave("../output/wide_both_diagnostics.pdf", h1, width = 12, height = 8)
```

# Wide dataset - standard normal with 10,000 dimensions
```{r}
# N <- 10000
# A <- diag(nrow=N)
# model <- stan_model("mvt_wide.stan")
# fit <- sampling(model, data=list(N=N), iter=400, chains=4)
# saveRDS(fit, "../output/mvt_wide_400.rds")
fit <- readRDS("../output/mvt_wide_400.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")
a_df <- tibble(samples_400_gbm=r_star_gbm,
               samples_400_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_400_summary.rds")
mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
```

1,000 samples
```{r}
# N <- 10000
# A <- diag(nrow=N)
# model <- stan_model("mvt_wide.stan")
# fit <- sampling(model, data=list(N=N), iter=1000, chains=4)
# saveRDS(fit, "../output/mvt_wide_1000.rds")

fit <- readRDS("../output/mvt_wide_1000.rds")
full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")

a_df <- readRDS("../output/mvt_wide_400_summary.rds")
a_df <- a_df %>% 
    mutate(samples_1000_gbm=r_star_gbm,
           samples_1000_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_summaries.rds")
mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
a_df <- readRDS("../output/mvt_wide_summaries.rds")
a_df_gbm <- a_df %>%
  select(contains("gbm")) %>% 
  rename(samples_400=samples_400_gbm,
         samples_1000=samples_1000_gbm) %>% 
  mutate(Classifier="C. GBM")
a_df_rf <- a_df %>%
  select(contains("rf")) %>% 
  rename(samples_400=samples_400_rf,
         samples_1000=samples_1000_rf) %>% 
  mutate(Classifier="D. RF")

g1 <- 
  a_df_gbm %>%
  bind_rows(a_df_rf) %>% 
  melt(id.vars="Classifier") %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("# of draws", labels=c("400", "1000")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=20, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.5, 0.5)) +
  facet_wrap(~Classifier)
ggsave("../output/very_wide_data.pdf", g, width = 12, height = 8)
```

```{r}
pdf("../output/mvt_wide_both.pdf", width = 12, height = 6)
multiplot(g, g1, cols = 1)
dev.off()
```